Corridor Project Proposal

### 2024-05-02 corridor_docs/masters_cmtee_proposal/eec_ms_proposal.Rmd
updating ms proposal
comments from corridor steering cmmmtee

---
title: "Corridor Project Proposal"
author: "Eric Escobar-Chena and Emilio M. Bruna"
date: 'updated: `r Sys.Date()`'
bibliography: "`r here::here('corridor_docs', 'eric-thesis.bib')`"

output: pdf_document
  # pdf_document: default
  # html_document:
  #   df_print: paged
  # word_document: default
geometry: margin=1in
mainfont: Times New Roman
fontsize: 12pt
indent: true
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \renewcommand{\headrulewidth}{0pt}
- \pagestyle{fancy}
- \fancyhead[L]{}
- \fancyhead[R]{Escobar-Chena, p. \thepage}
- \fancyfoot[C]{ }
- \fancyfoot[R]{updated `r Sys.Date()`}
- \AtBeginDocument{\let\maketitle\relax}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
- \usepackage{sectsty} \subsectionfont{\centering}
- \raggedright
- \usepackage{indentfirst}
- \setlength\parindent{24pt}
---

@@

All traps and markings will be removed at the study's conclusion.

[@WOS:000341406900011]

@@

<!-- \newpage -->

## Figures

@@

<!-- ![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%} -->

<!-- ![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%} -->

## References

>>>>

---
title: "Corridor Project Proposal"
author: "Eric Escobar-Chena and Emilio M. Bruna"
date: 'updated: `r Sys.Date()`'
bibliography: "`r here::here('corridor_docs', 'eric-thesis.bib')`"
csl: "`r here::here('corridor_docs', 'ecology.csl')`"
output: pdf_document
  # pdf_document: default
  # html_document:
  #   df_print: paged
  # word_document: default
geometry: margin=1in
mainfont: Times New Roman
fontsize: 12pt
indent: true
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \renewcommand{\headrulewidth}{0pt}
- \pagestyle{fancy}
- \fancyhead[L]{}
- \fancyhead[R]{Escobar-Chena, p. \thepage}
- \fancyfoot[C]{ }
- \fancyfoot[R]{updated `r Sys.Date()`}
- \AtBeginDocument{\let\maketitle\relax}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
- \usepackage{sectsty} \subsectionfont{\centering}
- \raggedright
- \usepackage{indentfirst}
- \setlength\parindent{24pt}
---

@@

All traps and markings will be removed at the study's conclusion. [@WOS:000341406900011]

@@

\newpage

# References

&nbsp;

::: {#refs custom-style="Bibliography"}
:::

\newpage

@@

![Arrangement of pitfall traps used to sample beetles in experimental blocks.](./images/plots1.png){fig-align="center" width=45% height=45%}

\newpage

![Locations and treatments for the dung removal experiment.](./images/plots3.png){fig-align="center" width=45% height=45%}

\newpage

![Design for the experimental assessment of beetle dispersal.](./images/plots2.png){fig-align="center" width=45% height=45%}

\newpage

## CORRIDOR STEERING COMMITTEE FEEDBACK

### Sampling in the matrix

**Doug:** What's the point of sampling in the matrix?  I think I know the answer.  I ask because it doesn't seem related to any of your predictions.

**Christopher Warneke:** I do like the idea of sampling the matrix to get a sense of what is going on with dung beetles there, so I am glad to see that in the proposal. When Carolyn Graham and I looked at foliar mites in the Corridor patches, we did look at the mites in the matrix too and found that mites were in higher abundance in the matrix. For our story, it was useful information for our Discussion, but it didn't matter too much for our study questions. But I could see it mattering for dung beetles, particularly if they are viewing the matrix and patches as essentially the same habitat type (though I doubt that is the case).

### Traps and Baits

**Doug: **I've trapped dung beetles in FL and learned that dung dries quickly in the sun and loses its attractiveness in the process.  My traps contained dung suspended from a modified coat-hanger in a bag made of panty hose.  Briefly soaking the bag in water 1-2 times/day kept the dung attractive.  I don't think I used 95% ethanol in the traps (maybe ethylene glycol?) because I was concerned about the odor of ethanol.  Finally, think about the type of dung.  I used cow dung but learned afterwards that it's not the best choice.  Pig dung is far better and probably easily available at UF.  There are papers on what type of dung works best (e.g., https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0073147

**Christopher Warneke:** I remember Sean Griffen (who worked with a megachild bee and movement between patches) that there were some challenges with number of bees found in the peripheral patches after being released in the center patch? I just glanced at the paper again [link here](https://nsojournals.onlinelibrary.wiley.com/doi/pdf/10.1111/ecog.05299) and it looks like it worked out better than my memory, but I worry about question 3 of the proposal and having enough recaps to be able to discuss movement rates (as it looks like is a common theme in feedback about the proposal). I don't know whether it would be helpful to check with Sean about things, but his website [is here](https://srgriffin.com/) if you want to reach out to him. Question 3 would be awesome if you had enough recaptures though!

**Lars: **Your preliminary surveys will help a lot, but what if you don’t catch much through your initial surveys. Would you look to ramp up the number of pitfalls you use in each patch (and would you have this capacity)?

I think there could be something interesting with the temporal dynamics, but it was less clear to me how critical this was compared to simply comparing among the different patch types. Do you know if there is a predominant season that you expect to catch the most beetles/when they might be most active dung foragers? If so, it might be worth considering ramping up sampling intensity during this time (and maybe sacrificing some of the temporal extent?). Just a suggestion – definitely feel free to ignore this!

### Dispersal

**Thomas:**  I think it is a good idea to assess species use of matrix both in dispersal trials and the richness sampling so you could control for this in species responses to the corridor. You could also include habitat specialist versus generalist users in the dispersal trials and potentially conduct an "inverse" of the Baum et al. (2004) corridor experiment, where instead of manipulating matrix directly, you use species that have different permeabilities or affinities for the matrix. The last things to think about with the dispersal trails are both the effect of the corridors on winds patterns and the prevailing wind direction for each block. I imagine olfaction is the primary way dung beetles locate dung sources, so this might be an important avenue with which either the corridors promote movement or corridor use gets obscured. One way to address this could be to either release in multiple unconnected patches or to sufficiently randomize across the blocks. If releases are conducted at the edge of patches, it might not require that only rectangular patches be used. I imagine the feasibility of this would depend on the numbers needed per release.

**Lars:**  Something to consider might also be to do multiple release distances from the bait, perhaps with different color markings (or something like that – I’m not sure how marking works), to work out beetle movement capacity. This could be preferable to having multiple bait distances, in case beetles are prone to stopping at the first bait they encounter. If you wanted to try out something like this, a recent large-ish clear cut could be good (bait in the middle, differing release distances radiating out from it) and I’m sure we could help locate one to work in.

### Beetles/Bycatch

**John Orrock: **In [this paper](https://link.springer.com/article/10.1007/s10980-011-9656-5) we report data from two years (6 total sampling sessions across 2001 and 2002) representing over 33,000 individuals from 58 genera.  The list of arthropods we captured is at the end (some scarabs, but we stopped at genus).  The pitfalls were relatively large (compared to centrifuge tubes and things we’ve used for other studies) so they collected a lot of things.  This paper uses pitfalls near the center of patches, but I also sampled near patch edges, in wings, and in corridors in 2022.  Those data were never published, but I think I just found a spreadsheet with them (over 13,000 individuals identified to genus).  That, said, the core-pitfall dataset in the attached paper is the one that is the most comprehensive.

### Assessing risk of plant introduction

**Christopher Warneke:** From a plant perspective, I wonder if it would potentially be possible to maybe plant some of the cow dung (or pig or whatever you go with) in the boneyard at station on site? For dung from ag animals, I suspect there won't be many germinants that come out from dung that gets moved by beetles, but if some is able to be planted at station, then Thomas/Benjamin/Eva could check on them and if any plants come up, that will help with the plant surveys within the patches if any plants pop up in the patches from the dung. That would just help us sort out unknown plants easier if we know they are accidental germinants from the cow/pig dung. Though I don't think this will really be a big issue, so it is probably also fine to not worry about it.

**Lars: **Christopher’s point about possibly introducing plants through dung is important (especially if this happens unequally across the patches, since one/block isn’t used). It’d be good to hear your thoughts on this possibility.
